Java Sorting Algorithms Comparison

### 2015-02-23 Programming Project B/paper.tex
added first two loop invariants to the paper

A merge sort can be implemented in a variety of languages using the pseudocode provided in Algorithm 2.
\newline
\textbf{Merge Sort Pre-Condition}: A is a non-empty array of a comparable data with a natural order.
\newline
\textbf{Merge Sort Post-Condition}: A' is a permutation of A (containing all the same elements) in strictly non-decreasing order.
\newline
\textbf{Merge Pre-Condition}: Left and right are both non-empty arrays of a comparable data type with a natural order in strictly non-decreasing order.
\newline
\textbf{Merge Post-Condition}: Combined has all the elements of both left and right in strictly non-decreasing order.
\begin{algorithm}
\caption {\textsc{Merge-Sort}(A)}
\label{algo:mergesort}
\begin{algorithmic}[1]
\Procedure{Merge-Sort}{A}
\If{$A.length < 2$}
\State{\Return{$A$}}
\EndIf
\State{$mid = A.length / 2$}
\For{$i = 1$ upto $mid$}
	\State{$left[left.length] = A[i]$}
\EndFor
\For{$i = mid$ upto $A.length$}
	\State{$right[right.length] = A[i]$}
\EndFor
\State{$left =$ Merge-Sort($left$)}
\State{$right =$ Merge-Sort($right$)}
\State{$A =$ Merge($left$, $right$)}
\State{\Return{$A$}}
\EndProcedure
\newline
\Procedure{Merge}{left, right}
\State{$var$ $i = 0$}
\State{$var$ $y = 0$}
\State{$var$ $x = 0$}
\While{$left.length$ $!=$ $i$ and $right.length$ $!=$ $y$}
	\If{$left[i] < right[y]$}
		\State{$combined[x] = left[i]$}
		\State{$i = i + 1$}
		\State{$x = x + 1$}
	\EndIf
	\If{$right[y] < left[i]$}
		\State{$combned[x] = right[y]$}
		\State{$y = y + 1$}
		\State{$x = x + 1$}
	\EndIf
\EndWhile
\While{$left.length$ $!=$ $i$}
	\State{$combined[x] = left[i]$}
	\State{$i = i + 1$}
	\State{$x = x + 1$}
\EndWhile
\While{$right.length$ $!=$ $y$}
	\State{$combined[x] = right[y]$}
	\State{$y = y + 1$}
	\State{$x = x + 1$}
\EndWhile
\State{\Return{$combined$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
\newline
%% TODO: Add pre and post conditions here
A heap sort can be implemented in a variety of languages using the pseudocode below in Algorithm 3.
\newline
\begin{algorithm}
\caption {\textsc{Heap-Sort}(A)}
\label{algo:heapsort}
\begin{algorithmic}[1]
\Procedure{Heapify}{A, i, total}
\State{$var$ $left = i * 2$}
\State{$var$ $right = left + 1$}
\State{$var$ $iPrime = i$}
\If{$left <= total$ and $A[left] > A[i]$}
	\State{$i = left$}
\EndIf
\If{$right <= total$ and $A[right] > A[i]$}
	\State{$i = right$}
\EndIf
\If{$i != iPrime$}
	\State{$var$ $temp = A[iPrime]$}
	\State{$A[iPrime] = A[i]$}
	\State{$A[i] = tmp$}
	\State{$A = heapify(A, i, total)$}
\EndIf
\State{\Return{$A$}}
\EndProcedure
\Procedure{Heap-Sort}{A}
\State{$var$ $size = A.length$}
\For{$var$ $i = size / 2$ downto $1$}
	\State{$A = heapify(A, i, size$}
\EndFor
\For{$var$ $i = size$ downto $1$}
	\State{$var$ $tmp = A[1]$}
	\State{$A[0] = A[i]$}
	\State{$A[i] = tmp$}
	\State{$size = size - 1$}
	\State{$A = heapify(A, 1, size)$}
\EndFor
\State{\Return{$A$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
%% END PROCEDURE

>>>>

A merge sort can be implemented in a variety of languages using the pseudocode provided in Algorithm 2.
\newline
\textbf{Merge Sort Pre-Condition}: A is a non-empty array of a comparable data with a natural order.
\newline
\textbf{Merge Sort Post-Condition}: A' is a permutation of A (containing all the same elements) in strictly non-decreasing order.
\newline
\textbf{Merge Pre-Condition}: Left and right are both non-empty arrays of a comparable data type with a natural order in strictly non-decreasing order.
\newline
\textbf{Merge Post-Condition}: Combined has all the elements of both left and right in strictly non-decreasing order.
\begin{algorithm}
\caption {\textsc{Merge-Sort}(A)}
\label{algo:mergesort}
\begin{algorithmic}[1]
\Procedure{Merge-Sort}{A}
\If{$A.length < 2$}
\State{\Return{$A$}}
\EndIf
\State{$mid = A.length / 2$}
\For{$i = 1$ upto $mid$}
	\State{$left[left.length] = A[i]$}
\EndFor
\For{$i = mid$ upto $A.length$}
	\State{$right[right.length] = A[i]$}
\EndFor
\State{$left =$ Merge-Sort($left$)}
\State{$right =$ Merge-Sort($right$)}
\State{$A =$ Merge($left$, $right$)}
\State{\Return{$A$}}
\EndProcedure
\newline
\Procedure{Merge}{left, right}
\State{$var$ $i = 0$}
\State{$var$ $y = 0$}
\State{$var$ $x = 0$}
\While{$left.length$ $!=$ $i$ and $right.length$ $!=$ $y$}
	\If{$left[i] < right[y]$}
		\State{$combined[x] = left[i]$}
		\State{$i = i + 1$}
		\State{$x = x + 1$}
	\EndIf
	\If{$right[y] < left[i]$}
		\State{$combned[x] = right[y]$}
		\State{$y = y + 1$}
		\State{$x = x + 1$}
	\EndIf
\EndWhile
\While{$left.length$ $!=$ $i$}
	\State{$combined[x] = left[i]$}
	\State{$i = i + 1$}
	\State{$x = x + 1$}
\EndWhile
\While{$right.length$ $!=$ $y$}
	\State{$combined[x] = right[y]$}
	\State{$y = y + 1$}
	\State{$x = x + 1$}
\EndWhile
\State{\Return{$combined$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
\newline
\textbf{Merge Sort First While Loop Invariant}: Left contains i elements, all of which can be found in A[0 ... i upto mid].
\newline
\textbf{Merge Sort First While Loop Invariant Initialization}: The invariant holds true vacuously because i is 0 and left contains no elements.
\newline
\textbf{Merge Sort First While Loop Invariant Maintenance}: The invariant holds true because i is incremented at the same rate that elements are added to left from the same i index value in A.
\newline
\textbf{Merge Sort First While Loop Invariant Termination}: The invariant holds true because i has been incremented at the same rate that elements have been added to left from the same i index value in A upto mid.
\newline
\newline
\textbf{Merge Sort Second While Loop Invariant}: Right contains i elements, all of which can be found in A[mid ... A.length].
\newline
\textbf{Merge Sort Second While Loop Invariant Initialization}: The invariant holds true vacuously because i is 0 and right contains no elements.
\newline
\textbf{Merge Sort Second While Loop Invariant Maintenance}: The invariant holds true because i is incremented at the same rate that elements are added to right from the same i index value in A.
\newline
\textbf{Merge Sort Second While Loop Invariant Termination}: The invariant holds true because i has been incremented at the same rate that elements have been added to right from the same i index value in A upto to A.length.
\newline

%% TODO: Add pre and post conditions here
A heap sort can be implemented in a variety of languages using the pseudocode below in Algorithm 3.
\newline
\begin{algorithm}
\caption {\textsc{Heap-Sort}(A)}
\label{algo:heapsort}
\begin{algorithmic}[1]
\Procedure{Heapify}{A, i, total}
\State{$var$ $left = i * 2$}
\State{$var$ $right = left + 1$}
\State{$var$ $iPrime = i$}
\If{$left <= total$ and $A[left] > A[i]$}
	\State{$i = left$}
\EndIf
\If{$right <= total$ and $A[right] > A[i]$}
	\State{$i = right$}
\EndIf
\If{$i != iPrime$}
	\State{$var$ $temp = A[iPrime]$}
	\State{$A[iPrime] = A[i]$}
	\State{$A[i] = tmp$}
	\State{$A = heapify(A, i, total)$}
\EndIf
\State{\Return{$A$}}
\EndProcedure
\Procedure{Heap-Sort}{A}
\State{$var$ $size = A.length$}
\For{$var$ $i = size / 2$ downto $1$}
	\State{$A = heapify(A, i, size$}
\EndFor
\For{$var$ $i = size$ downto $1$}
	\State{$var$ $tmp = A[1]$}
	\State{$A[0] = A[i]$}
	\State{$A[i] = tmp$}
	\State{$size = size - 1$}
	\State{$A = heapify(A, 1, size)$}
\EndFor
\State{\Return{$A$}}
\EndProcedure
\end{algorithmic}
\end{algorithm}
%% END PROCEDURE
